### 2019-03-29 paper/arxiv/main.tex
submitted article

\documentclass[aps,prb,11pt]{revtex4-1}

\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bm}

\usepackage{hyperref}
\newcommand{\ie}{i.\,e.~}
\newcommand{\Ie}{I.\,e.~}
\newcommand{\eg}{e.\,g.~}
\newcommand{\Eg}{E.\,g.~}
% \newcommand{\e}{\mathrm{e}}
\newcommand{\ic}{i}

\newcommand*\conj[1]{
  \vbox{
  \hrule height 0.3ptword
  \kern0.5ex
  \hbox{
  \kern-0.4em
  \ifmmode#1\else\ensuremath{#1}\fi
   \kern-0.em
  }
 }
}
\newcommand{\B}{\boldsymbol}
% \newcommand{\tens}[1]{\B{#1}}
% \newcommand{\tens}[1]{\underline{\underline{#1}}}
\newcommand{\tens}[1]{\B{#1}}
\newcommand{\re}{\mathrm{Re}}
\newcommand{\im}{\mathrm{Im}}
\newcommand{\grad}{\B{\mathrm{\nabla}}}
\renewcommand{\div}{\B{\mathrm{\nabla\cdotp}}}
\newcommand{\ddroit}{\mathrm{d}}
\newcommand{\epsf}{\varepsilon^{\rm f}}
\newcommand{\epsftens}{\tens{\varepsilon}^{\rm f}}
\newcommand{\epstens}{\tens{\varepsilon}}
\newcommand{\epsd}{\varepsilon^{\rm d}}
\newcommand{\epsvac}{\varepsilon_{0}}
\newcommand{\epshom}{\tilde{\epstens}}
\newcommand{\fig}[1]{Fig.~(\ref{#1})}
\newcommand{\equ}[1]{Eq.~(\ref{#1})}

\newcommand{\xitens}{\tens{\xi}}
\newcommand{\xihom}{\tilde{\xitens}}

%
% \usepackage[outdir=./]{epstopdf}
% \graphicspath{{./}{../../data/figures/}}

\begin{document}

%%%% Article title to be placed here
\title{Homogenization of periodic and random ferroelectric-dielectric composites}
\author{Benjamin Vial and Yang Hao}

%%%%%%%%% Insert author address here
\affiliation{School of Engineering and Computer Science, Queen Mary, University of London, London, E1 4NS, United Kingdom}

% %%%% Subject entries to be placed here %%%%
% \subject{Electromagnetism, Computer modelling and simulation, Applied mathematics}
%
% %%%% Keyword entries to be placed here %%%%
% \keywords{Homogenization, Ferroelectrics, Metamaterials , Tunability}
%
% %%%% Insert corresponding author and its email address}
% \corres{Benjamin Vial\\
%  \email{b.vial@qmul.ac.uk}}

%%%%%%%%%%%%%%% End of first page %%%%%%%%%%%%%%%%%%%%%

%

%%%% Abstract text to be placed here %%%%%%%%%%%%
\begin{abstract}
 We investigate the homogenized permittivity of ferroelectric-dielectric mixtures under
 a static electric field. A refined model is used to take into account the coupling
 between the electrostatic problem and the electric field dependent permittivity of the
 ferroelectric material. Periodic and random structures in two dimensions are investigated and
 we study the effective permittivity, losses, electrically induced anisotropy and tunability
 of those metamaterials.
\end{abstract}
%%%%%%%%%%%%%%%%%%%%%%%%%%%

\maketitle

 \section{Introduction}
 Ferroelectric materials play a crucial role in reconfigurable
 microwave devices, with typical applications including antenna beam steering,
 phase shifters, tunable power splitters, filters, voltage controlled oscillators and
 matching networks \cite{tagantsev_ferroelectric_2018}. Both bulk ceramics and thin films have
 been employed to design frequency agile components \cite{vendik_ferroelectric_1999, lancaster_thin-film_1998,xi_oxide_2000} and metamaterials \cite{hand_frequency_2008, zhao_experimental_2008}.
 The main reason of using
 ferroelectric materials is their strong dependence of their permittivity $\varepsilon$
 on an applied electric field $E$, which is measured by their tunability defined as $n = \varepsilon(0)/\varepsilon(E)$.
 The key requirements for antenna and microwave applications are large tunability and low losses.
 These two characteristics are correlated and one has to find a trade-off for optimal device performance, which can be quantified by the so called commutation quality factor
 $K = (n -1)^2/(n\, \tan\delta(0)\,\tan\delta(E))$, where $\tan\delta$ is the loss tangent.\\
 These materials have usually high permittivity values,
 often leading to slow response time and impedance mismatch, which can be an issue in some practical
 applications. Thus it has been considered to mix ferroelectric ceramics to low-index and
 low-loss non-tunable dielectrics in order to reduce both permittivity

\noindent values and losses.
The effective parameters of those composites have been investigated \cite{sherman_ferroelectric-dielectric_2006, jylha_tunability_2008, sherman_tunability_2004, astafiev_can_2003} and it has been found that the
permittivity can be greatly reduced while losses are much less sensitive to the dielectric phase addition, and in some situations lead to a small increase of the tunability of the mixture. In the context of porous ferroelectrics, the homogenized properties strongly depends on the size and morphology of the pores \cite{okazaki_effects_1973,stanculescu_study_2015}. In addition, the influence
of local field enhancement in the tunability has been shown to be of particular importance \cite{padurariu_tailoring_2012,padurariu_field-dependent_2012,cazacu_tunable_2013}.
\\
This study investigates the effective permittivity of composites made of dielectric inclusions in a ferroelectric matrix
by using a two-scale convergence method \cite{allaire_homogenization_1992, guenneau_homogenization_2000}.
The originality lies in the fact that a fully coupling model is employed to
calculate the electrostatic field distribution when a uniform biasing field is
applied on the structures, which will result in a local modification of the permittivity
in the ferroelectric phase due to the microstructure. As compared to a simple uncoupled model where the
ferroelectric phase is only modified through the biasing field,
the resulting effective permittivity, dielectric losses, tunability and
anisotropy significantly differ. As compared to earlier sudies in the litterature \cite{padurariu_tailoring_2012,padurariu_field-dependent_2012}, we account for
the non-linear coupling beyond the first iteration and use two-scale
convergence homogenization analysis to obtain the effective parameters at higher frequencies, instead of a capacitance-based model.
We first study metamaterials consisting of a squarre array of parallel dielectric rods with circular cross section in a ferroelectric host, and then investigate the effect of random distribution of those rods within the unit cell.

%#########################################################################################

\section{Theory and numerical model}

We consider a composite made of a ferroelectric material with anisotropic
permittivity $\epsftens(\B E)$
that is dependent on an applied electric field $\B E$,
and a non tunable dielectric of permittivity $\epsd$, which are both non-magnetic.
The structures under study are invariant along the $z$ direction, which leads to the standard decomposition
of the wave equation in the transverse electric case (TE, electric field parallel
to the direction of invariance) and the
transverse magnetic case (TM, magnetic field parallel to the direction
of invariance).
A uniform biasing field is applied in order to be able to tune the effective permittivity. Modelling homogenized properties of this type of mixtures can be done by assuming
that the electric field distribution is uniform throughout the sample,
so that the study of the tunability is
essentially achieved by changing the value of the properties in the ferroelectric phase and computing the effective permittivity of the composite.
We refer this approach as to the uncoupled model in the following.
However, a more accurate description is to take into account the change of the
electric field by the microstructure, if any. We therefore need to solve an
electrostatic equation to find the field distribution within the material, but its solution
depends on the permittivities of both materials, and the permittivity in the
ferroelectric phase depends on this induced electric field: this leads to a strongly coupled problem.\\

% -------------------------------------------------------------------------------
\subsection{Permittivity model\label{permmodel}}

We use barium strontium titanate (BST) as our ferroelectric material. Measurements
have been carried out at electrostatics and at microwave frequencies, and the normalized permittivity value
as a function of biasing field are reported on \fig{fig1}.\\
To describe the permittivity, we make use of the Landau potential
given by $F(P,E) = F_0 +  a P^2/2 + b P^4/4 + cP^6/6 - EP$, where $E$ is
the applied electric field and $P$ is the polarization \cite{landau_electrodynamics_2013, zhou_dielectric_2008}. Variations of the
permittivity with the temperature can be taken into account through the
coefficients $a$, $b$ and $c$, but we assume we are working at a constant
room temperature. We further assume that the material is not subject to any stress, so that the variation
of permittivity due to mechanical constraints is irrelevant.
The equation of state $$\frac{\partial F (P, E)}{\partial P}   = a P_0 + b P_0^3 + c P_0^5 - E = 0$$ gives the
dependence of the polarization on the applied electric field,
with $P_0$ being the equilibrium polarization.
Along the direction of a uniform applied electric field, the relative permittivity is given by:
\begin{equation}
 \epsf(E) = \left[\frac{\partial^2 F (P, E)}{\partial P^2} \right]^{-1} = \frac{\epsf(0)}{1 + \alpha P_0^2 + \beta P_0^4},
 \label{eq_epsf}
\end{equation}
where $\epsf(0)=1/a$, $\alpha=3b/a$ and $\beta=5c/a$. The fitting parameters are given in Table \ref{table_params_fit}.
As the norm of the field increases, the permittivity decreases with a characteristic
bell curve typical for a ferroelectric material in its paraelectric state.
\begin{figure}[!t]
 \centering
 \includegraphics[width=0.75\columnwidth]{epsilon_fit}
 \caption{Variation of the ferroelectric permittivity as a function of the
  applied electric field (dots: measurements, dashed lines: fit to
  formula~(\ref{eq_epsf})), for the static case (green) and at microwave frequencies
  (orange, $f=3.8$ GHz). The fitting parameters are given in Table \ref{table_params_fit}.}
 \label{fig1}
\end{figure}
Furthermore, assuming the crystalline principal axes of the ferroelectric material
are oriented in the coordinate directions, and that the diagonal components of the permittivity
tensor are only function of the corresponding bias electric field components \cite{krowne_anisotropic_2002}, we have:
\begin{equation}
 \epsftens (\B E) =
 \left(
 \begin{matrix}
   \epsf_{xx}(E_x) & 0               & 0 \cr
   0               & \epsf_{yy}(E_y) & 0 \cr
   0               & 0               & \epsf_{zz}(E_z) \cr

  \end{matrix}
 \right)
 \label{eq_epsftens}
\end{equation}
where each of the diagonal components have the functional form
given by \equ{eq_epsf}. Note that we will use the static values of permittivity
for the electrostatic modelling, while we are interrested in the homogenized values
of permittivity at microwaves.\\
\begin{table}
 \caption{Fitting parameters to model (\ref{eq_epsf}) for the measured permittivity values as a function
  of applied electrc field shown on \fig{fig1}. }
 \label{table_params_fit}
 \begin{tabular}{llll}
  \hline
  case        & $\epsf(0)$ & $\alpha$ ($\rm \mu m^2/V^2$) & $\beta$ ($\rm \mu m^4/V^4$) \\ \hline
  static      & 3050       & 0.120                        & 0.024                       \\
  $f=3.8$ GHz & 165        & 0.240                        & 0.079                       \\\hline
 \end{tabular}
\end{table}

% -------------------------------------------------------------------------------
\subsection{Electrostatic model}
The composites under study are made of two materials, thus their permittivity
is represented by a piecewise defined tensor $\epstens(\B r, \B E)$ which is
equal to $\epsftens(\B E(\B r))$ in the ferroelectric phase and ${\rm diag}(\epsd)$
in the dielectric phase.
In the following, we consider two different cases for the biasing field.
Because of the form~(\ref{eq_epsftens}) assumed for the ferroelectric permittivity
tensor, $\varepsilon_{zz}$ will not be changing for a field in the plane orthogonal
to the $z$ axis. This is the only component
being relevant for TE polarization, so we consider in this case a uniform biasing
electric field applied along the direction of invariance $\B E_0 = E_{0} \B e_z$.
On the other hand,
the in-plane components of $\epsftens$ are tuned by $E_x$ and $E_y$, therefore,
without loss of generality,
we consider a uniform applied electric field directed along the $x$ axis
$\B E_0 = E_{0} \B e_x$ for the TM polarization case.
To calculate the total electric field in the material, one
has to solve the electrostatic equation for the potential $V$:
\begin{equation}
 \div (\epstens \grad V) = 0
 \label{eq_elstat}
\end{equation}
Note that for the TE case, the solution is trivial since the structures
are invariant along $z$, so that the electric field is equal to the uniform biasing field, and
we will thus not study it in the following.
However in the TM case, the situation is much more complex: this is a coupled problem since the
electric field $\B E=-\grad V$ derived from the
solution of \equ{eq_elstat} depends on the permittivity distribution, which
itself depends on the electric field.
The coupled system formed
of Eqs.~(\ref{eq_epsftens}) and (\ref{eq_elstat}) is solved iteratively until there
is convergence on the norm of the electric field.
Here we would like to emphasise that the permittivity in the ferroelectric material, although
uniform initially, is spatially varying due to the non-uniform distribution
of the total electric field.\\

% -------------------------------------------------------------------------------
\subsection{Homogenization}
The effective permittivity for TM polarization is calculated
using two scale homogenization technique \cite{allaire_homogenization_1992, guenneau_homogenization_2000}.
For this purpose, one has to find the solutions
$\psi_j$ of two annex problems $\mathcal P_j$, $j=\{1, 2\}$:
\begin{equation}
 \div \left[ \xitens \grad(\psi_j + r_j) \right] = 0,
 \label{eq_hom_annex}
\end{equation}
where $\B r = (x, y)^{\rm T}$ is the position vector in the $xy$ plane and
$\xitens=\epstens^{\rm T}/ {\rm det }(\epstens)$.
The homogenized tensor $\xihom$ is obtained with:
\begin{equation}
 \xihom = \langle \xitens \rangle + \B \phi,
 \label{eq_hom}
\end{equation}
where $\langle . \rangle$ denotes the mean value over the unit cell.
The elements of the matrix $\B \phi$ represent correction terms and
are given by $\B \phi_{ij} = \langle \xitens \grad \psi_i \rangle_j$.
Finally the effective permittivity tensor can be calculated using $\epshom=\xihom^{\rm T}/ {\rm det }(\xihom)$.\\
Note that the TE case, which we shall not study here as no coupling happens, is trivial since
the homogenized permittivity is simply the average of the permittivity in the unit cell:
$\epshom = \langle \epstens \rangle$.

% -------------------------------------------------------------------------------
\subsection{Numerical setup}
Equations~(\ref{eq_elstat}) and (\ref{eq_hom_annex}) are solved with a Finite Element
Method using the open source packages Gmsh \cite{geuzaine_gmsh:_2009} and GetDP \cite{dular_general_1998}.
In both cases we use a square unit cell $\Omega$ of length $d$ with periodic boundary
conditions along $x$ and $y$. Second order Lagrange elements are used and the
solution is computed with a direct solver (MUMPS \cite{amestoy_fully_2001}).
% The source code and
% data necessary to reproduce the results reported in this paper are freely available
% online.

%#########################################################################################

\section{Numerical results}
In the following numerical results, the dielectric phase is supposed to be
lossless and non dispersive with $\epsd = 3$ while the ferroelectric material follows the
permittivity described in section~\ref{permmodel} and has a constant loss
tangent $\tan \delta^{\rm f} = 10^{-2}$.

% -------------------------------------------------------------------------------
\subsection{Two dimensional periodic metamaterial}
\begin{figure}[!t]
 \centering
 \includegraphics[width=1\textwidth]{convergence_per}
 \caption{Convergence of the coupled problem.
  Real part (a) and loss tangent (b) of the components of the homogenized
  permittivity tensor as a function of iteration step $i$. The distribution of
  the normalized electric field (colour map: magnitude in logarithmic scale,
  lines: equipotential contours) and of the
  $xx$ component of the permittivity tensor are shown for $i=1$
  (c and d) and $i=12$ (e and f).
 }
 \label{conv2D}
\end{figure}
Lets us now consider a periodc square array of infinitely long dielectric rods of circular cross section
of radius $r$ embedded in a ferroelectric matrix.\\
We first study the convergence of the coupled problem on the particular case with dielectric
filling fraction $f=\pi r^2/d^2=0.5$ and $E_0=2$MV/m. Figures \ref{conv2D}(a) and \ref{conv2D}(b) show the
convergence of the real part and loss tangent of the components of the homogenized
permittivity tensor, respectively. The $yy$ components converge quickly
and are almost unaffected by the coupling process whereas the
$xx$ components change substantially from the initial conditions.
This is due to the effect of the redistribution
of the electrostatic field within the unit cell (see Figs.~(\ref{conv2D}.c) and (\ref{conv2D}.d)),
where the $x$ component of the electric field is still much stronger
than the $y$ component, even if spatially varying in the ferroelectric medium.
At equilibrium, the electric field is concentrated close to the $y$ axis in between two neighbouring
rods. This in turn affects the permittivity distribution (see Figs.~(\ref{conv2D}.e) and (\ref{conv2D}.f)),
and the homogenized properties of the composite.\\
We computed the effective parameters of these metamaterial structures for different
radii of the rods and studied their behaviour when subjected to an external
electrostatic field (see \fig{eff_par_2D_TM}). The results of our coupled
model differ significantly from the uncoupled one. Increasing the dielectric fraction
lowers the effective permittivity while the losses are slightly reduced but much less sensitive.
Due to the inhomogeneous redistribution of the permittivity over the ferroelectric domain, the
overall tunability changes. In the case studied here, taking into account
the coupling leads to an effective tunability increase with
higher dielectric concentration, and that is larger than the tunability
 of bulk ferroelectric. Two concurrent effects are at stake here: on the one hand
the dilution of ferroelectric makes the composite less tunable, but on the other hand,
the rearrangement of the electrostatic field surrounding the inclusion and its
concentration in some region will cause a higher permittivity change locally.
The relative strength of those phenomena is governed by the shape of the inclusion and its permittivity
and so it is envisioned that the performance of the composites might be enhanced by engineering
their microstructure.
The geometry of the unit cell is symmetric so the homogenized material is
isotropic when no field is applied.
But when the sample is biased, the permittivity distribution becomes asymmetric due
to the inhomogeneity of the electric field, thus making the effective material properties anisotropic.
This geometric effect is added to the anisotropy arising from the material properties of the ferroelectric
phase itself, and depending on the topology and permittivity of the rods, one effect would be predominant.
In the case studied here, the equilibrium permittivity distribution varies strongly along the bias
 direction and much less orthogonally to it, which adds anisotropy by diminishing the effective
 permittivity in the $x$ direction. This is local field induced effect is what makes the anisotropy stronger
 in our coupled model compared to the uncoupled one (cf. Fig.~(\ref{eff_par_2D_TM}.d)).
Those subtle phenomena can only be rigorously taken into account by employing a coupling formalism
and are responsible for the difference observed when compared to a simple uncoupled model.\\

\begin{figure}[!t]
 \centering
 \includegraphics[width=0.9\textwidth]{effective_params_per}
 \caption{Effective parameters of the 2D metamaterials as a function of the
  applied electric field for various filling fraction of dielectric.
  (a): normalized permittivity, (b): normalized loss tangent, (c): normalized tunability and
  (d): anisotropy factor. The solid lines correspond to the coupled model and
  the dashed lines to the uncoupled model.}
 \label{eff_par_2D_TM}
\end{figure}

% -------------------------------------------------------------------------------
\subsection{Random case}

We finally study the effect of random particle distribution on the effective parameters of the
composites. This is an important point as fabrication of randomly dispersed
inclusions is much more easy from a technological perspective. For each filling fraction of the dielectric,
we generated 21 numerical samples with inclusions of circular cross section of average radius
$r=d/20$ that can vary by $\pm 30\%$. Their centre is chosen randomly and the
rods are allowed to overlap. An example of distribution for $f=0.5$ is given on \fig{randmatepsi}.
The effective material properties are plotted on \fig{eff_par_2Drand_TM}.
Similarly to the periodic case, the permittivity decreases with increasing dilution of
ferroelectric, but for identical filling fraction,
the permittivity is lower as compared to the periodic array, and the smaller the dielectric concentration the larger
is the difference. Losses decrease as well and the reduction is substantially larger
than the periodic case, with higher variation from sample to sample as $f$ increases.
The effective tunability is on average smaller than that in the periodic case, and
for low biasing fields and for some particular samples can be greater than the bulk tunability. However,
at higher applied electric fields, normalized tunability becomes smaller than unity and
is reduced as one adds more dielectric. For comparison, the homogenized
parameters are plotted on \fig{eff_par_2Drand_TM_uncpl} in the case where the coupling is
neglected. One can see that the coupled and uncoupled models give
similar results for the tunability whereas the losses are still smaller for the coupled case
at higher fields.\\
The redistribution of electric field, permittivity and convergence of
the effective parameters are displayed in \fig{conv_random}. The effect of
disorder plays an important role here: the electrostatic field gets concentrated
in between neighbouring inclusions and the smaller the gap the higher the field, hence
a greater local permittivity change. In addition, even if the distribution of particle is
random, one expects that the anisotropy due to geometry would cancel for a sufficiently large number
of particles (which is the case as the mean anisotropy factor is close to $1$ when no
bias field is applied). However, the anisotropy due to ferroelectric properties is
important in this case as well, as both the $x$ and $y$ components of the electrostatic field
are playing a role.
Because of the relative positions of the rods, both $\varepsilon_{xx}$ and $\varepsilon_{yy}$ are affected
by the coupling, so that the anisotropy factor for higher fields is reduced as compared to the periodic case.
However even if there is a substantial variability from sample to sample, on average, the anisotropy factor
decreases with increasing dielectric concentration.

\begin{figure}[!t]
 \centering
 \includegraphics[width=0.9\textwidth]{effpar_rand_cpl.png}
 \caption{Effective parameters of the random 2D mixtures as a function of the
  applied electric field for various filling fraction of dielectric, when the
  coupling is taken into account.
  (a): normalized permittivity, (b): normalized loss tangent, (c): normalized tunability and
  (d): anisotropy factor. The solid lines represent the average values
  over the 21 samples and the lighter error bands show a confidence interval corresponding to
  the standard deviation.}
 \label{eff_par_2Drand_TM}
\end{figure}
%
%
%
\begin{figure}[!t]
 \centering
 \includegraphics[width=1\textwidth]{convergence_rand}
 \caption{Convergence of the coupled problem in the random case
  for one sample.
  Real part (a) and loss tangent (b) of the components of the homogenized
  permittivity tensor as a function of iteration step $i$. The distribution of
  the normalized electric field (colour map: magnitude in logarithmic scale,
  lines: equipotential contours) and of the $xx$ component of the permittivity tensor are shown for $i=1$
  (c and d) and $i=12$ (e and f).
 }
 \label{conv_random}
\end{figure}

%#########################################################################################
%

\section{Conclusion}

We have studied the homogenized properties of dielectric/ferroelectric mixtures
using a rigorous model that take into account the coupling between the electrostatic
field distribution and the field dependant ferroelectric permittivity tensor. After
convergence of the coupled problem, the effective permittivity tensor is calculated using
two scale convergence homogenization theory.
The results obtained by this model differ significantly from a simple assumption that
the permittivity of the ferroelectric respond just to the uniform biasing field.
We have considered both periodic and random arrays
of dielectric rods in a ferroelectric matrix in 2D, and studied their effective properties
for TM polarization as a function of dielectric concentration and bias field.
Importantly, adding more low index and low loss dielectric allows to decrease
the overall permittivity significantly and slightly lower the losses.
For the periodic case, the tunability is higher than the bulk due to local field enhancement, whereas
this effect is strongly suppressed when disorder is introduced. The assymetric redistribution of the permittivity
induce en effective anisotropy that is added to the one arising purely from the ferroelectric material.
The properties of the composites are affected by multiple factors:
geometry and the spatially dependent electric field that will induce locally a tunable, anisotropic
response in the ferroelectric phase depending on its amplitude and direction.
This suggest that the performances of the composites
may be enhanced by distributing the two phases in an optimal way to get high
tunability and low losses. Further work in that direction is needed as well as
extending this study to 3D media.
Finally, because the permittivity of the dielectric is much smaller than the ferroelectric one,
it would be of great interest to use high contrast homogenization theory
\cite{bouchitte_homogenization_2004, cherednichenko_homogenization_2015} to
study this kind of mixtures.
This would reveal the frequency dependant artificial magnetism due to "micro-resonances"
in the high index phase and potentially lead to composites with tunable effective permeability.

\section*{Additional information}

\vskip6pt

\noindent{The codes necessary to reproduce the results in this article are freely
available online at this address: \href{https://www.github.com/benvial/ferromtm}{https://www.github.com/benvial/ferromtm}.}\\
{Benjamin Vial wrote devised the research, implemented
the codes and wrote the paper. Yang Hao supervised the research and
co-authored the manuscript.}\\
{The authors declare no competing interrests.}\\
{This work was funded by the Engineering and Physical Sciences Research
Council (EPSRC), UK, under a grant (EP/P005578/1) 'Adaptive Tools for
Electromagnetics and Materials Modelling to Bridge the Gap between
Design and Manufacturing (AOTOMAT)'.}\\
{The authors would like to thanks Henry Giddens for performing the
measurements of ferroelectric permittivity used in this paper.}\\

\appendix

\section{Random permittivity distribution and uncoupled effective parameters}

\begin{figure}[h!]
 \centering
 \includegraphics[width=0.7\columnwidth]{random_samples_f_50percent}
 \caption{Permittivity distribution of the numerical samples used for $f=0.5$. Dark
  colour indicates the ferroelectric material while light colour represents the
  dielectric inclusions.}
 \label{randmatepsi}
\end{figure}

\begin{figure}[h!]
 \centering
 \includegraphics[width=0.8\textwidth]{effpar_rand_uncpl.png}
 \caption{Effective parameters of the random 2D mixtures as a function of the
  applied electric field for various filling fraction of dielectric, when the
  coupling is neglected.
  (a): normalized permittivity, (b): normalized loss tangent, (c): normalized tunability and
  (d): anisotropy factor. The solid lines represent the average values
  over the 21 samples and the lighter error bands show a confidence interval corresponding to
  the standard deviation.}
 \label{eff_par_2Drand_TM_uncpl}
\end{figure}
%

% \disclaimer{Insert disclaimer text here.}

%%%%%%%%%% Insert bibliography here %%%%%%%%%%%%%%

\vskip2pc

\begin{thebibliography}{99}

\bibitem{tagantsev_ferroelectric_2018}
Tagantsev AK, Sherman VO, Astafiev KF, Venkatesh J, Setter N. 2018
  Ferroelectric {Materials} for {Microwave} {Tunable} {Applications}. {\em J.
  Electroceram.} \textbf{11}, 5--66.

\bibitem{vendik_ferroelectric_1999}
Vendik O, Hollmann E, Kozyrev A, Prudan A. 1999  Ferroelectric tuning of planar
  and bulk microwave devices. {\em Journal of Superconductivity} \textbf{12},
  325--338.

\bibitem{lancaster_thin-film_1998}
Lancaster M, Powell J, Porch A. 1998  Thin-film ferroelectric microwave
  devices. {\em Superconductor Science and Technology} \textbf{11}, 1323.

\bibitem{xi_oxide_2000}
Xi X, Li HC, Si W, Sirenko A, Akimov I, Fox J, Clark A, Hao J. 2000  Oxide thin
  films for tunable microwave devices. {\em Journal of Electroceramics}
  \textbf{4}, 393--405.

\bibitem{hand_frequency_2008}
Hand TH, Cummer SA. 2008  Frequency tunable electromagnetic metamaterial using
  ferroelectric loaded split rings. {\em Journal of Applied Physics}
  \textbf{103}, 066105.

\bibitem{zhao_experimental_2008}
Zhao H, Kang L, Zhou J, Zhao Q, Li L, Peng L, Bai Y. 2008  Experimental
  demonstration of tunable negative phase velocity and negative refraction in a
  ferromagnetic/ferroelectric composite metamaterial. {\em Applied Physics
  Letters} \textbf{93}, 201106.

\bibitem{sherman_ferroelectric-dielectric_2006}
Sherman VO, Tagantsev AK, Setter N, Iddles D, Price T. 2006
  Ferroelectric-dielectric tunable composites. {\em Journal of Applied Physics}
  \textbf{99}, 074104.

\bibitem{jylha_tunability_2008}
Jylha L, Sihvola AH. 2008  Tunability of granular ferroelectric dielectric
  composites. {\em Progress in Electromagnetics Research} \textbf{78},
  189--207.

\bibitem{sherman_tunability_2004}
Sherman VO, Tagantsev AK, Setter N. 2004  Tunability and loss of the
  ferroelectric-dielectric composites. In {\em 14th {IEEE} {International}
  {Symposium} on {Applications} of {Ferroelectrics}, 2004. {ISAF}-04. 2004} pp.
  33--38.

\bibitem{astafiev_can_2003}
Astafiev KF, Sherman VO, Tagantsev AK, Setter N. 2003  Can the addition of a
  dielectric improve the figure of merit of a tunable material?. {\em Journal
  of the European Ceramic Society} \textbf{23}, 2381 -- 2386.

\bibitem{okazaki_effects_1973}
Okazaki K, Nagata K. 1973  Effects of {Grain} {Size} and {Porosity} on
  {Electrical} and {Optical} {Properties} of {PLZT} {Ceramics}. {\em Journal of
  the American Ceramic Society} \textbf{56}, 82--86.

\bibitem{stanculescu_study_2015}
Stanculescu R, Ciomaga CE, Padurariu L, Galizia P, Horchidan N, Capiani C,
  Galassi C, Mitoseriu L. 2015  Study of the role of porosity on the functional
  properties of ({Ba},{Sr}){TiO}3 ceramics. {\em Journal of Alloys and
  Compounds} \textbf{643}, 79--87.

\bibitem{padurariu_tailoring_2012}
Padurariu L, Curecheriu L, Galassi C, Mitoseriu L. 2012a  Tailoring non-linear
  dielectric properties by local field engineering in anisotropic porous
  ferroelectric structures. {\em Appl. Phys. Lett.} \textbf{100}, 252905.

\bibitem{padurariu_field-dependent_2012}
Padurariu L, Curecheriu L, Buscaglia V, Mitoseriu L. 2012b  Field-dependent
  permittivity in nanostructured {BaTiO}$_3$ ceramics: {Modeling} and
  experimental verification. {\em Phys. Rev. B} \textbf{85}, 224111.

\bibitem{cazacu_tunable_2013}
Cazacu A, Curecheriu L, Neagu A, Padurariu L, Cernescu A, Lisiecki I, Mitoseriu
  L. 2013  Tunable gold-chitosan nanocomposites by local field engineering.
  {\em Appl. Phys. Lett.} \textbf{102}, 222903.

\bibitem{allaire_homogenization_1992}
Allaire G. 1992  Homogenization and {Two}-{Scale} {Convergence}. {\em SIAM
  Journal on Mathematical Analysis} \textbf{23}, 1482--1518.

\bibitem{guenneau_homogenization_2000}
Guenneau S, Zolla F. 2000  Homogenization of {Three}-{Dimensional} {Finite}
  {Photonic} {Crystals}. {\em Journal of Electromagnetic Waves and
  Applications} \textbf{14}, 529--530.

\bibitem{landau_electrodynamics_2013}
Landau LD, Bell J, Kearsley M, Pitaevskii L, Lifshitz E, Sykes J. 2013 {\em
  Electrodynamics of continuous media} vol.~8.
elsevier.

\bibitem{zhou_dielectric_2008}
Zhou K, Boggs SA, Ramprasad R, Aindow M, Erkey C, Alpay SP. 2008  Dielectric
  response and tunability of a dielectric-paraelectric composite. {\em Applied
  Physics Letters} \textbf{93}, 102908.

\bibitem{krowne_anisotropic_2002}
Krowne CM, Daniel M, Kirchoefer SW, Pond JA. 2002  Anisotropic permittivity and
  attenuation extraction from propagation constant measurements using an
  anisotropic full-wave {Green}'s function solver for coplanar ferroelectric
  thin-film devices. {\em IEEE Transactions on Microwave Theory and Techniques}
  \textbf{50}, 537--548.

\bibitem{geuzaine_gmsh:_2009}
Geuzaine C, Remacle JF. 2009  Gmsh: {A} 3-{D} finite element mesh generator
  with built-in pre- and post-processing facilities. {\em International Journal
  for Numerical Methods in Engineering} \textbf{79}, 1309--1331.

\bibitem{dular_general_1998}
Dular P, Geuzaine C, Henrotte F, Legros W. 1998  A general environment for the
  treatment of discrete problems and its application to the finite element
  method. {\em IEEE Transactions on Magnetics} \textbf{34}, 3395--3398.

\bibitem{amestoy_fully_2001}
Amestoy PR, Duff IS, Koster J, L'Excellent JY. 2001  A {Fully} {Asynchronous}
  {Multifrontal} {Solver} {Using} {Distributed} {Dynamic} {Scheduling}. {\em
  SIAM Journal on Matrix Analysis and Applications} \textbf{23}, 15--41.

\bibitem{bouchitte_homogenization_2004}
Bouchitté G, Felbacq D. 2004  Homogenization near resonances and artificial
  magnetism from dielectrics. {\em Comptes Rendus Mathématique} \textbf{339},
  377 -- 382.

\bibitem{cherednichenko_homogenization_2015}
Cherednichenko K, Cooper S. 2015  Homogenization of the system of high-contrast
  {Maxwell} equations. {\em Mathematika} \textbf{61}, 475--500.

\end{thebibliography}

% \bibliographystyle{RS}
% \bibliography{biblio}

\end{document}