### 2019-06-10 poster/main_poster.tex
poster

\documentclass[final,12pt]{beamer} % use beamer
\include{preamble}

\begin{document}
\color{fresnelblack}
\date{\today}
\title{Effective parameters of ferroelectric dielectric mixtures}
\subtitle{}
\author{B. Vial and Y. Hao }

\begin{frame}

%%%%%%%%%%%%%%%%%%Title, authors and affiliations   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{textblock*}{80cm}(2cm,2cm)
{\hfill \blackfont\fontsize{1.9cm}{1em}\selectfont \textcolor{fresnelorange}{\MakeUppercase{\inserttitle}}\\}\vspace*{2em}
{\hfill \textcolor{white}{\Large\insertauthor}\\}
\vspace*{16pt}
{\hfill \itshape\large\textcolor{white}{
School of Electronic Engineering \& Computer Science, Queen Mary, University of London}}
\end{textblock*}

%%%%%%%%%%%  ABSTRACT %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{textblock*}{79.5cm}(2.5cm,16.2cm)
% \textcolor{fresneldarkblue}{\rule{\columnwidth}{5pt}}\\
%  {\color{fresneldarkblue}\bfseries Abstract:}
{\color{white}{ \blackfont  \bfseries \large \uppercase{Abstract}\\} \vspace{0.5em} 
  \large \itshape
  We investigate the homogenized parameters of ferroelectric-dielectric composites under a static electric field. A numerical model that takes into account the coupling
  between the electrostatic problem and the electric field dependent permittivity of the
  ferroelectric material is used. Metamaterials consisting of periodic and random arrays of rods
  are considered for transverse electric polarization case
  and we study their effective permittivity, losses, electrically induced anisotropy
  and tunability by a two scale convergence homogenization method.}
\end{textblock*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%% Left column  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% 
% 
%%%%%%%%%%%  BOX1 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{textblock*}{38.5cm}(2.5cm,29cm)

\begin{posterblock}{Method}{benchcolor1}

\begin{columns}[t,totalwidth=\columnwidth]
\begin{column}{.55\columnwidth}

  Ferroelectric materials play a crucial role in reconfigurable
  microwave devices, with typical applications including antenna beam steering,
  phase shifters, tunable power splitters, filters, voltage controlled oscillators and
  matching networks \cite{tagantsev_ferroelectric_2018}.

\mysubtitle{Ferroelectric permittivity}

\begin{equation}
 \epsf(E) = \left[\frac{\partial^2 F (P, E)}{\partial P^2} \right]^{-1} = \frac{\epsf(0)}{1 + \alpha P_0^2 + \beta P_0^4},
 \label{eq_epsf}
\end{equation}

\mysubtitle{Normalization of modes}

\mysubtitle{Expansion of the coupled modes}

\end{column}
\begin{column}{.40\columnwidth}
% \vspace*{2cm}
% \begin{figure}
% \centering
% \input{./figures/schemafig1}
%  \end{figure}

\end{column}
\end{columns}

\end{posterblock}
\end{textblock*}
% 
% % 
% %%%%%%%%%%%%% BOX 3 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% 
% \begin{textblock*}{38.5cm}(2.5cm,72cm)
% \begin{posterblock}{Results}{benchcolor1}
%  \mysubtitle{Dielectric rod dimmer}\vspace{-1.5cm}
% 
% \vspace*{0.7cm}
% TE polarization $\B E=E_z\B z$, $\varepsilon_a=\varepsilon_b=10-0.01\ic$, $\varepsilon_h=1$, 
% $\mu_a=\mu_b=\mu_h=1$, $L_x=10^{-2}\,$m, $L_y=3L_x$, gap $d$, $\omega_0=2\pi c/L_y$.
% 
% \begin{columns}[b,totalwidth=\columnwidth]
% \begin{column}{.49\columnwidth}
%  \begin{figure}
%  \centering
%  \includegraphics[width=\columnwidth]{./figures/figd2}
%  \caption{Real (top) and imaginary (bottom) parts of the eigenfrequencies for the first four 
%  coupled modes as a function of normalized gap $d/L_x$. Circles: coupled modes (full wave direct computation), 
%  solid lines: 1 mode coupling model.}
% \end{figure}
% \end{column}
% \begin{column}{.49\columnwidth}
% \vspace*{1cm}\\
% \begin{figure}
%  \centering
%  \includegraphics[width=0.5\columnwidth]{./figures/singlerod1}
%   \includegraphics[width=0.5\columnwidth]{./figures/singlerod2}
%  \caption{Eigenfield $\re(E_z)$ for the modes of the isolated rod.}
% \end{figure}
% \begin{figure}
%  \centering
%  \includegraphics[width=0.5\columnwidth]{./figures/test1}
%   \includegraphics[width=0.5\columnwidth]{./figures/test2}\\
%    \includegraphics[width=0.5\columnwidth]{./figures/test3}
%   \includegraphics[width=0.5\columnwidth]{./figures/test4}\\
%  \caption{Eigenfield $\re(E_z)$ for the coupled modes, $d/L_x=2.1$.}
% \end{figure}
% 
% \end{column}
% \end{columns}
% \vspace*{1cm}
% Very good agreement is obtained with this simple one mode model. \\
% The study of the phase difference between the coefficients $A_i$ and $B_i$ 
% indicates an in-phase (resp. out of phase) hybridization mechanism 
% for modes $C1$ and $C3$ (resp. $C2$ and $C4$).
% 
% \end{posterblock}
% 
% \end{textblock*}
% % 
% % 
% % % 
% % % %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% % % 
% % %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% % %%%%%%%%%%% Right column  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% % %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% % 
% % %%%%%%%%%%%%%%%%%%  BOX %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% % 
% \begin{textblock*}{38.5cm}(43.5cm,29cm)
% \begin{posterblock}{Results}{benchcolor1}
% 
%  \mysubtitle{Breaking the symmetry}
% Shift $e$ along $y$, $d/L_x=1$.
% 
% \begin{columns}[b,totalwidth=\columnwidth]
% \begin{column}{.49\columnwidth}
% 
% \begin{figure}
% \includegraphics[width=\columnwidth]{./figures/figfne}
%  \caption{Real (top) and imaginary (bottom) parts of the eigenfrequencies for the first four 
%  coupled modes as a function of normalized shift $e/L_y$. Circles: coupled modes (full wave direct computation), 
%  solid lines: 4 modes coupling model.}
%  \end{figure}
% \end{column}
% \begin{column}{.49\columnwidth}
% 
%  \begin{figure}
%   \includegraphics[width=\columnwidth]{./figures/AN}\\
%   \includegraphics[width=0.5\columnwidth]{./figures/etest3}
%   \includegraphics[width=0.5\columnwidth]{./figures/etest4}
%  \caption{Top: norm of coupling coefficient $A$ as a function of 
%  normalized shift $e/L_y$. Bottom: coupled modes for $e/L_y=0.2$.}
%  \end{figure}
% \end{column}
% \end{columns}
% 
% \vspace*{1cm}
% Good agreement is obtained between full wave calculations and our model. \\
% The shift breaks the symmetry of the coupling mechanism, as revealed by 
% the study of the coefficients $A_i$ and $B_i$. For modes $C3$ and $C4$, 
% both modes $A1$ and $A2$ are significantly coupled when $e\neq0$.
% \end{posterblock}
% \end{textblock*}
% % 
% % 
% %%%%%%%%%%%%%%%%%%  BOX %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% 
% \begin{textblock*}{38.5cm}(43.5cm,68.3cm)
% \begin{posterblock}{Conclusion and outlook}{benchcolor1}
% We developped a model for the coupling of quasi-normal modes in meta-molecules, allowing 
% for a fast computation of coupled complex eigenfrequencies and QNMs. The validity of 
% our approach has been illustrated with an example of a dielectric nanorod dimmer. It 
% is found that shifting one of the resonator perpendicular to the dimmer axis induce a 
% symmetry breaking of the coupling mechanism.\\
% Future research directions includes:
% \begin{itemize}
%    \item model for TM polarization
%    \item include magnetic material	
%    \item include dispersion \cite{QNM3}
%    \item more than two coupled resonators
%     \item 3D model
%   \end{itemize}
% 
% \end{posterblock}
% 
% \end{textblock*}
% % % 
% % % 
% %%%%%%%%%%%%%%%%%%%%%  Reference box %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% 
\begin{textblock*}{38.5cm}(43.5cm,86.cm)
\myblocktitle{References}{fresnelgreenblue}
\color{fresnelblack}
\printbibliography
\end{textblock*}

%%%%%%%%%%%%%%%%%%%%%  Acknowledgments box %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{textblock*}{38.5cm}(43.5cm,105.cm)
\myblocktitle{Acknowledgments}{fresnelgreenblue}
\color{fresnelblack}
This work was funded by the Engineering and Physical
Sciences Research Council (EPSRC), UK under a
Programme Grant (EP/I034548/1) ‘‘The Quest for
Ultimate Electromagnetics using Spatial Transformations
(QUEST).’’
\end{textblock*}
%

\begin{textblock*}{70cm}(12cm,113cm)

{\hfill\semiboldfont \bfseries \color{fresnelorange} \Large CONTACT \& LINKS}\\
\vspace*{0.5em}
\color{white}
\hfill Corresponding author: Benjamin Vial |~\textcolor{fresnelorange}{\href{mailto:b.vial@qmul.ac.uk}{\faEnvelope~b.vial@qmul.ac.uk}}\\
\hfill QUEST |~\textcolor{fresnelorange}{\href{http://www.quest-spatial-transformation.org}{\faExternalLink~www.quest-spatial-transformation.org}}\\
\hfill Antennas \& Electromagnetics Research Group |~\textcolor{fresnelorange}{\href{http://antennas.eecs.qmul.ac.uk}{\faExternalLink~antennas.eecs.qmul.ac.uk}}

\end{textblock*}
%% LOGOS %%%%%%%%%%%

\begin{textblock*}{5.5cm}(2.5cm,114cm)
 \includegraphics[width=\columnwidth]{./logos/logo_epsrc.png}
\end{textblock*}

\begin{textblock*}{6cm}(10cm,114.5cm)
 \includegraphics[width=\columnwidth]{./logos/logo_qmul.png}
\end{textblock*}
% 
% \begin{textblock*}{10.8cm}(17.6cm,114cm)
%  \includegraphics[width=\columnwidth]{./logos/logo_qmul.png}%{logo_black.png}
% \end{textblock*}
% 
% \begin{textblock*}{8cm}(30.cm,114cm)
%  \includegraphics[width=\columnwidth]{./logos/logo_exeter.png}
% \end{textblock*}
% 
% \begin{textblock*}{8.7cm}(39.5cm,114cm)
%  \includegraphics[width=\columnwidth]{./logos/logo_oxford.png}
% \end{textblock*}

\end{frame}

\end{document}